Study of Intra- and Inter-Job Interference for Improving Allocation on HPC Systems

### 2015-10-12 submission.tex
draft version V5

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{url}
\usepackage{xcolor}
\usepackage{amsmath}
%\usepackage{subfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage[labelfont=bf]{caption}
\usepackage[font=bf,labelsep=space]{caption}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Study of Intra/Inter-Application Interference for Smart Job Allocation on HPC system}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

@@

\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
Illinois Institute of Technology,
Chicago, Illinois, USA 60616\\
{xyang56, zzhou}@hawk.iit.edu, lan@iit.edu}

@@

It has been widely recoginzed that network contension between concurrently running jobs in HPC system is one of the primary causes of performance variability. Optimizing job allocation and avoiding network sharing have been proved to be effective for alleviating network contension and performance degradation. In this work, we propose a new job communication pattern aware allocation strategy which serve as a critical module in our future HPC batch scheduler. The novelty of our new strategy is that it makes allocation decision with not only system topology and resource availability information, but also with job's communication pattern information. The information scheduler gets about job communication pattern can help it to decide whether the job’s performance relies heavily on the network or not, and what the potential inter/intra-job interference would be. Thus, the scheduler can pick the preferable resource allocation based on specific job's communication pattern and preserve the locality of allocation in a better way. Using traces collected from three typical parallel applications of DOE, we validate the effectiveness of our new desgin. The idea presented in this paper is widely applicable: HPC parallel application has specific dominant communication pattern, allocating resource with communication pattern awareness can preserve locality in a bettwer way on systems with differnt network topologies.

@@

The scale of supercomputer keeps growing from petascale to exascale, to accomodate the the demand of insatiable computing power from scientific research areas. Production systems contain hundreds of thousands of processors serve as irreplaceable research vehicle for scientific problems with increasing size and complexity. Supercomputers employed in a shared way to accomodate many jobs running concurrently\ref{fig: mira activity} , which will also improve system utilization. These jobs share the system infrastructure such as network, I/O bandwidth, which will inevitably cause contention over the shared resource. As supercomputer continues to evolve, these shared resources are tend to become the bottleneck for performance.

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{God.}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Bad.}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Indication of multiple jobs running on Cray system. Many jobs runninng concurrently on the system. Different jobs represented by specific colors.}
   \label{fig: overview}
\end{figure}

@@

One of the most prominant problem comes into researcher's attention is the network contention between concurrently running jobs. The network sharing between concurrently running jobs cause communication variability, which reslut in jobs running slower and longer waiting time for results. The delay of currently running jobs will increase the queueing time of the following submitted jobs, thus leads to low system throughput, low utilization and high energy cost. The network sharing can even cause interference between neighbouring jobs, abort the the running job unexpectedly, or even worse, wrong science results.

This adverse effect of network sharing can be mitigated by providing the job with better node allocation. The batch scheduler is responsible for assigning job with the amount of node it required upon submission. Currently two scheduling strategies are commonly used on torus-connected systems. One is so called partition based systems, where the scheduler assigns each user job a compact and contiguous set of computing nodes. IBM Blue Gene series systems fall into this category. This strategy is in favor of application’s performance by providing it with isolated partition and exclusive network connections, thus reducing network contention caused by concurrently running jobs sharing network bandwidth. However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation(when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to poor system performance (e.g., low system utilization and long waiting time for job). The other is non-contiguous allocation system, where free nodes are assigned to user job no matter whether they are contiguous or not. Cray XT/XE series systems fall into this category. Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization. Nevertheless, it introduces other problems such as scattering application processes all over the system and causing both inter-job contention and intra-job contention. The non-contiguous node allocation can make inter-process communication less efficient and cause network contention among concurrently running jobs, thereby resulting in poor job performance especially for those communication-intensive jobs.

Partition-based allocation achieves good job performance by sacrificing system performance (e.g., poor system utiliza- tion), whereas non-contiguous allocation can result in better system performance but could severely degrade performance of user jobs (e.g, prolonged wait-time and run-time). Supercomputers keep growing in size, its network become more sophisticated then ever before and applications also increase its complexity. How to provide application with better allocation that can balance job performance with system performance become a prominant problem need a solution before HPC enters era of exasacle computing. In this work, we propose a guideline for future batch scheduling system, a smart way to make allocation with the awareness of job's communication pattern.

Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it use partition-based or non-contigous policy, means it chooses the benefit of either system or job over the other. Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information.

Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}. Job's communication trace and performance analysis can be get by using such tool at runtime.

The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions. This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{hpdc2015-ornl}. The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. Job's communication traces study would be extreme helpful to leadership computing facility too. The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness. We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}. In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.

In this work, we present a guideline for future batch scheduler to make allocation with awareness about job's communication pattern. Rather than choose blindly between partition-base or non-contigous allocation, the new allocation mechanism make allocation based on job's communication pattern. With detail analysis about application's communication pattern, we can get a clear view of its communication topology graph, data transfer between processes, and decide which processes are tightly connected to which and conduct intensive communication. The "local community" consists of such tightly coupled processes shuld get compact alloction. Then, the allocation mechanism should make allocations that only needs to guarantee the locality of the subset of the nodes where those "local community" should reside on.

We first pick three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct deep study about their communication patterns. We identify the "local community" in these applications and assign them allocation accrodingly. The advantage of making allocation is obvious. First, compared with partition-based policy, our communication pattern aware policy is more flexible, it doesn't require the system to provide a big partition that can accomodate the whole application, just small set of compact nodes enough for all the "local community" in the application. On the other hand, our communication pattern aware allocation policy won't destroy the locality of application's communication pattern. On the contrary, it provides small compact node sest for the the "local communities" in the application will preserve the communication locality in the maximum extent.

We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture, from Argonne National Lab as research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus network. CODES enables the exploration of simulating different HPC networks with high fidelity\cite{Jason-2011}. We extend CODES with a Job Mapping API that support different allocation policies on torus network.

We also conduct a case study with the results we got from communication pattern aware allocation policy evaluation with real trace from production system.\textcolor{red}{Not Sure About this part!!!}

The rest of this paper orgainzed as follows. Section\ref{sec:application study} gives a detailed study of the three applications from DOE Desgin Forward Project. Section\ref{sec:simulation} talks about the simulation platform and Section

@@

A parallel application usuall conforms to a combination of several basic communication patterns. At its different execution phases, the application's communication behavior may follow different basic patterns repectively. When we look into the data flow during the application execution, most parallel application start with broadcast operation to distribute the data from root process to other processes, followed by a series of computation and communication that comforms to certain pattern, which is usually the dominant part of application's execution. And before the application come to completion, all the working process will return their results to the root dirctly or hierarchically. In this work, we focus on the dominant communication pattern because that's where the application spend most of its running time. We believe making optimized allocation regarding application's dominant communication pattern would greatly benefit both the application and the system. In this section, we analysis two applications whose dominant communication patterns are most prevalent in scientific computation.

There are many profiling tools available to capture the communication behavior from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip}. They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.

In this work, we provide detailed analysis about the communication pattern of two DOE MiniApps. MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications\cite{miniapp}. Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures. We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter. Their communication patterns can be representative for most MiniApps listed on the DOE Design Forward Webpage\cite{design forward webpage}. The communication pattern figures we present here are generated with the IPM\cite{ipm} data from \cite{design forward webpage}.

\subsection{AMG}
\label{sec:amg}
AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}. The dominant communication pattern is this regional communication with decreasing message size for different parts of the multigrid v-cycle.

@@

Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG. The blue line shows that total data amount transferred from each rank, while the red shows the amount of received data and green shows the amount of send data. The first thing we can spot from the figure is that for each rank, the amount of data sent and received are basically the same, which also well explain the symmetry of the application's dominant communication pattern. And the second thing we can find about AMG is that the data transfer amount also has a symmetric pattern. Rank 0 has the same amout of data transfer as Rank 26, Rank 1 as of Rank 25, etc.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_pip}
   \caption{Rank-Rank communication topology of AMG with 27 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}.   }
   \label{fig: amg communication topology}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_data_transfer}
   \caption{The amount of data each rank transferred in AMG. }
   \label{fig: amg data trans}
\end{figure}

@@

The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000. Nek5000\cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. The MiniApp of Nek5000, Crystal Router demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_pip}
   \caption{Rank-Rank communication topology of CrystalRouter with 100 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
   \label{fig: crystalrouter100 communication topology}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_data_transfer}
   \caption{The amount of data each rank transferred in CrystalRouter. }
   \label{fig: cr data trans}
\end{figure}

As we can see from figure \ref{fig: crystalrouter100 communication topology}, every 5-rank is a compact community, where data transfer is very intensive. Beside the "5-rank community", there is "pair-wise" global data transfer in Crystal Router. Obviously, there is data transfer alone the diagnal. When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a multi-stage manner. The same pattern can be observed from the global data transfer. Our observation about the dominant communication pattern of Crystal Router is the multi-stage "5-rank community" and "pair-wise" global data transfer happen recursively as the scale of the application increase.

@@

\subsection{MultiGrid}
\label{sec:multigrid}

Differe from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes. MultiGrid conforms to 3D lattice communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle. It is widely used for structured grid physics packages.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg_pip}
   \caption{Rank-Rank communication topology of MultiGrid with 125 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
   \label{fig: multigrid125 communication topology}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg_data_transfer}
   \caption{The amount of data each rank transferred in MultiGrid. }
   \label{fig: cr data trans}
\end{figure}

\section{Research Vehicle}
\label{sec:simulation}

A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity\cite{Jason-2011}\cite{mubarak-sc2012}. CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers\cite{ross}. CODES support both torus and dragonfly network with high fidelity flit-level simulation. In this work, we only use torus network to show the impact of node allocation to job with different dominant communication patterns. CODES has this network workload component that capable of conducting trace-driven simulation. It can take real MPI application trace generated by SST DUMPI\cite{sst} to drive CODES network models.

We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies. The new API can help us to explore the impact of different allocation strategies to job with specific communication pattern. We have choosen two publicly available HPC network traces provided by the Design Forward Program\cite{design forward webpage}. The detailed analysis of these two traces are present in Section\ref{sec:amg} and \ref{sec:crystalrouter}.

The torus network performance is determined by its dimensionality and link bandwidth. As the increase of the dimensionality of torus netowrk, so does the number of links connected with each node. The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application. Figure \ref{fig: dimenstudy} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2). The bandwidth between nodes is 2GiB/s in one direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus. As we can see from these figures, communication time of both applications decrease as the dimensionality of the network increase. The increased aggregated bandwidth of each node can accelerate the transfer of data. This can justify the fidelity and consistency of the torus network model. The same pattern can be found in Figure \ref{fig:bandwidthstudy}, as the bandwidth increase from 2GiB/s, to 4GiB/s and 8GiB/s in the 5D torus network, the data transfer time can be greatly reduced.

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/dimenstudy/avgtime}
        \caption{Normalized Average Time.}
        \label{fig:1sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/dimenstudy/maxtime}
        \caption{Normalized Max Time.}
        \label{fig:1sub2}
    \end{subfigure}%
   \caption{Normalized time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
   \label{fig: dimenstudy}
\end{figure}

>>>>

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{url}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{textcomp}
%\usepackage{subfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage[labelfont=bf]{caption}
\usepackage[font=bf,labelsep=space]{caption}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Study of Intra- and Inter-Job Interference for Improving Allocation on HPC Systems}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

@@

\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
Illinois Institute of Technology,
Chicago, Illinois, USA 60616\\
\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}

@@

It has been widely recognized that network contention between concurrently running jobs in HPC system is one of the primary causes of performance variability. Optimizing job allocation and avoiding network sharing are proven techniques to alleviating network contention and performance degradation. In this work, we propose a novel job communication pattern aware allocation strategy which serve as a critical module in our future HPC batch scheduler. The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern communication pattern in addition to system topology and resource availability information. The scheduler uses job communication patterns information to decide whether a job’s performance relies heavily on the network or not, and what the potential inter/intra-job interference would be for a given allocation. Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern. Using traces have been collected from three representative parallel applications, we validate the effectiveness of our new design. The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on systems with a variety of network topologies.

@@

The scale of supercomputers keeps growing to accommodate the the demand for computing power required by scientific research areas. Supercomputers contain hundreds of thousands of nodes, and serve as irreplaceable research vehicle for scientific problems with increasing size and complexity. Supercomputers are employed as shared resource to accommodate many jobs running concurrently\ref{fig: overview} and maintain high system utilization. These jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over the shared resources. As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{Contiguous}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Non-contiguous}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Indication of multiple jobs running concurrently with different allocations. Different job represented by specific color. a) shows the effect of contiguous allocation, which would alleviate the inter-job interference. b) shows non-contiguous allocation, which definitely introduce both intra and inter-job interference. }
   \label{fig: overview}
\end{figure}

@@

One of the most prominent  problems is network contention between concurrently running jobs. The network sharing between concurrently running jobs causes communication variability that results in jobs running slower\cite{abhinav-sc13}. The delay of currently running jobs increases the queueing time of the following submitted jobs, thus leading to low system throughput, low utilization\cite{jose-ipdps15}.

This adverse effect of network sharing can be mitigated by providing jobs with better node allocations. The batch scheduler is responsible for assigning job the amount of nodes it requires. Currently, two allocation strategies are commonly used on torus-connected systems. One is so called partition based allocation, where the scheduler assigns each user job a compact and contiguous set of computing nodes. IBM Blue Gene series systems use this strategy. This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs. However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to low system utilization than otherwise possible. The other strategy is non-contiguous allocation, where free nodes are assigned to user jobs whether they are contiguous or not. Cray XT/XE series systems use this strategy. Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization. However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention. The non-contiguous node allocation can significantly reduce job performance, especially for those communication-intensive jobs.

%\textcolor{red}{
%Partition-based allocation achieves good job performance by sacrificing system utilization, whereas non-contiguous allocation can result in better system utilization but could severely degrade performance of user jobs. What is needed is a more flexible strategy that provides allocations that minimize intra/inter-job contention but allow greater system utilization. In this work, we propose a guideline for future batch scheduling systems, a smart way to allocate resources with the awareness of jobs communication patterns.
%
%Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other. Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information. 
%
%Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}. Job's communication trace and performance analysis can be get by using such tool at runtime. 
%
%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions. This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{hpdc2015-ornl}. The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. Job's communication traces study would be extreme helpful to leadership computing facility too. The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness. We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}. In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
%}

%=========come back and work from here=============
Rather than choose blindly between partition-based or non-contiguous allocation, the new allocation strategy choses resources based on a job's communication pattern. With detail analysis of application's communication patterns, a clear view of its communication topology is obtained shows which processes are tightly connected and conduct intensive communication. The ``local community" consists of such tightly coupled processes should get a compact allocation. Then, the allocation strategy only needs to guarantee the locality of the subset of the nodes where those ``local communities" would reside.

In this work, we focus on the study of impact of different allocation strategies on intra/inter-job interference on Torus network. Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance. On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Lab and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.

We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct deep study about their communication patterns. We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus networks. We analysis each application's performance when it running on torus networks with different dimensionality and interconnected bandwidth. Then, we extend CODES with a Job Mapping API that support different allocation policies on torus network. Different allocation strategies on torus network are studied with the help of this new Job Mapping API. We analysis the intra/inter-job interference by simulating those three applications running exclusively or concurrently on torus network with different allocation strategies. According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy. We believe the idea presented in this work would be the guideline for the design of future HPC batch job scheduler and resource management module.

The rest of this paper organized as follows. Section \ref{sec:application study} gives a detailed study of the three applications from DOE Design Forward Project. Section \ref{sec:codes} talks about CODEs as research vehicle for our work. Section \ref{sec:torus} shows the performance analysis of three applications on different torus networks. Section \ref{sec:interference} provides detailed analysis about the intra/inter-job interference between three applications on torus network. Section \ref{sec:discussion} talks about the idea of smart allocation strategy with job communication pattern awareness. Section \ref{sec:related_work} talks about the related work in the area and conclusion will be presented in Section \ref{sec:conclusion}.

@@

%A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. \textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution. Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc. In this work, we focus on the locality of application's rank-to-rank communication topology graph.

A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. Collective communication is often used for data distribution and collection among processes. Nearest Neighbor communication is commonly seen in lots of structured and unstructured mesh physics packages. There are many profiling tools available to capture communication patterns from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip}. They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.

In this work, we provide detailed analysis about the communication pattern of three DOE MiniApps. MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications\cite{miniapp}. Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures. We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter. Their representative communication patterns that are commonly seen in the HPC system workload. The communication pattern figures we present here are generated with the IPM\cite{ipm} data from \cite{design forward webpage}.

\subsection{AMG}
\label{sec:amg}

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
        \caption{Communication Topology}
        \label{fig: amg communication topology}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
        \caption{Data Volume}
        \label{fig: amg data trans}
    \end{subfigure}
    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in AMG.}
\end{figure}

AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}. The dominant communication pattern is this regional communication with decreasing message size for different parts of the multigrid v-cycle.

@@

Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG. The blue line shows that total data amount transferred from each rank, while the red shows the amount of received data and green shows the amount of send data. The first thing we can spot from the figure is that for each rank, the amount of data sent and received are basically the same, which also well explain the symmetry of the application's dominant communication pattern. And the second thing we can find about AMG is that the data transfer amount also has a symmetric pattern. Rank 0 has the same amount of data transfer as Rank 26, Rank 1 as of Rank 25, etc.

AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor. Applications with such similar dominant communication pattern are like PARTISN, SNAP\cite{design forward webpage}.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_pip}
        \caption{Communication Topology}
        \label{fig: cr communication topology}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
        \caption{Data Volume}
        \label{fig: cr data trans}
    \end{subfigure}
    \caption{CrystalRouter. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in CrystalRouter. }
\end{figure}

The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000. Nek5000\cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. The communication kernel of Nek5000, CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process. \textcolor{yellow}{The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.}

As we can see from Figure \ref{fig: cr communication topology}, every 5-rank is a compact community, where data transfer is very intensive. Beside the "5-rank community", there is "pair-wise" global data transfer in CrystalRouter. Obviously, there is data transfer alone the diagnal. \textcolor{yellow}{This is because messages would be buffered on their node of origin until a synchronization point was reached when a single system call sent every message to its destination in one pass.} When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a hierarchical manner. The same pattern can be observed from the global data transfer. Our observation about the dominant communication pattern of CrystalRouter is the multi-stage "5-rank community" and "pair-wise" global data transfer happen hierarchically as the scale of the application increase.

@@

CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication.

\subsection{MultiGrid}
\label{sec:multigrid}
Different from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes. MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle. It is widely used for structured grid physics packages. Figure\ref{fig: mg communication topology} shows the communication topology of MultiGrid and Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid.

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_pip}
        \caption{Communication Topology}
        \label{fig: mg communication topology}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
        \caption{Data Volume}
        \label{fig: mg data trans}
    \end{subfigure}
    \caption{MultiGrid. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in MultiGrid.}
\end{figure}

MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).

%\begin{figure*}[t!]
%    \centering
%    \begin{subfigure}[t]{0.32\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
%        \caption{AMG}
%        \label{fig: amg communication topology}
%    \end{subfigure}%
%    ~ 
%    \begin{subfigure}[t]{0.32\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_pip}
%        \caption{CrystalRouter}
%        \label{fig: cr communication topology}
%    \end{subfigure}
%    ~
%    \begin{subfigure}[t]{0.32\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_pip}
%        \caption{MultiGrid}
%        \label{fig: mg communication topology}
%    \end{subfigure}
%    \caption{Caption place holder}
%\end{figure*}
%
%\begin{figure*}[t!]
%    \centering
%    \begin{subfigure}[t]{0.32\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
%        \caption{AMG DATA}
%        \label{fig: amg data trans}
%    \end{subfigure}
%    ~
%    \begin{subfigure}[t]{0.32\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
%        \caption{CR DATA}
%        \label{fig: cr data trans}
%    \end{subfigure}
%    ~
%    \begin{subfigure}[t]{0.32\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
%        \caption{MG DATA}
%        \label{fig: mg data trans}
%    \end{subfigure}
%    \caption{Caption place holder}
%\end{figure*}

%We summarize the benchmarks and MiniApps on DOE design forward webpage\cite{design forward webpage}, as shown in Table \ref{tab: app summary}. According to their dominant communication patterns, there are three categories, namely Nearest Neighbor, All-to-All and 3D Lattice. The three applications studies in this work are representative for different categories.
